CF Registry Analysis - ROC

### 2025-09-15 Christine Chan/CF Registry/cf_registry_ROC_analysis.qmd
file paths and formatting

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(ggplot2)
library(gtsummary)
library(gt)
library(pROC)
library(ggrepel)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
  "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/GitHub",
  "Windows" = "C:/Users/Tim/GitHub",
  "Linux" = "/home/tim/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
set.seed(1017)
```

```{r data cleaning}
# Import data
encounter <- read.csv("./Data_Raw/CFF19_encountersMerged_Del1.csv",
  na.strings = c("", "N/A")
)
# Demographic data
demo <- read.csv("./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
  na.strings = c("", "N/A")
)
# Race
races <- c(
  "White", "Black or African American", "American Indian or Alaska Native",
  "Asian", "Native Hawaiian or Other Pacific Islander", "Some other race"
)
demo$Race <- apply(demo, 1, function(r) {
  w <- which(r[paste0("Race", 1:6)] == 1)
  if (length(w) == 0) {
    "Unknown"
  } else if (length(w) > 1) {
    "More than one race"
  } else {
    races[w]
  }
})
demo$Race <- factor(demo$Race,
  levels = c(
    "White", "Black or African American", "American Indian or Alaska Native",
    "Asian", "More than one race", "Native Hawaiian or Other Pacific Islander",
    "Some other race"
  ),
  labels = c(
    "White", "Black or African American", "Other race", "Other race",
    "More than one race", "Other race", "Other race"
  )
)
demo$Hispanicrace <- factor(demo$Hispanicrace,
  levels = 1:2,
  labels = c("Yes", "No")
)
# F508
demo$F508 <- factor(demo$F508,
  levels = 1:3,
  labels = c("F508del/F508del", "F508del/Other", "Other/Other")
)
# Age at year end
annual <- read.csv("./Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")
)
annual <- annual %>% rename(reviewyear = ReviewYear)
# Merge
encounter <- left_join(
  encounter,
  demo %>% select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
  by = join_by(eDWID)
)
encounter <- left_join(
  encounter,
  annual %>% select(eDWID, reviewyear, Age_YrEnd),
  by = join_by(eDWID, reviewyear)
)
# Limit to those with two hour OGTT and HbA1c at the same encounter
# Also remove years before 2010 and those under age 10
roc_df <- encounter %>%
  arrange(eDWID, reviewyear, encounternum) %>%
  select(
    eDWID, reviewyear, Age_YrEnd, Race, Hispanicrace, Gender, F508, MutClass,
    twohour_bloodglucose, hgba1c
  ) %>%
  filter(reviewyear >= 2010, Age_YrEnd >= 10, hgba1c >= 3.5 & hgba1c <= 15) %>%
  drop_na() %>%
  group_by(eDWID) %>%
  filter(row_number() == n()) %>%
  ungroup() %>%
  mutate(
    `OGTT 2-hour` = cut(twohour_bloodglucose,
      breaks = c(-Inf, 140, 200, Inf),
      labels = c("<140", "[140,200)", ">=200")
    ),
    `Age Group` = cut(Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
  )
# Metrics to report
metrics <- c(
  "thresh", "specificity", "sensitivity", "accuracy", "tn", "tp", "fn",
  "fp", "npv", "ppv"
)
```

>>>>

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(ggplot2)
library(gtsummary)
library(gt)
library(pROC)
library(ggrepel)
home_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Christine Chan/CF Registry",
    "Windows" = "C:/Users/tim/OneDrive - The University of Colorado Denver/Vigers/CF/Christine Chan/CF Registry",
    "Linux" = "/home/tim/OneDrive/Vigers/CF/Christine Chan/CF Registry"
)
github_dir <- switch(
    Sys.info()["sysname"],
    "Darwin" = "/Users/timvigers/GitHub",
    "Windows" = "C:/Users/Tim/GitHub",
    "Linux" = "/home/tim/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
set.seed(1017)
```

```{r data cleaning}
# Import data
encounter <- read.csv(
    "./Data_Raw/CFF19_encountersMerged_Del1.csv",
    na.strings = c("", "N/A")
)
# Demographic data
demo <- read.csv(
    "./Data_Raw/CFF19_DemogCFDiag_Del1.csv",
    na.strings = c("", "N/A")
)
# Race
races <- c(
    "White",
    "Black or African American",
    "American Indian or Alaska Native",
    "Asian",
    "Native Hawaiian or Other Pacific Islander",
    "Some other race"
)
demo$Race <- apply(demo, 1, function(r) {
    w <- which(r[paste0("Race", 1:6)] == 1)
    if (length(w) == 0) {
        "Unknown"
    } else if (length(w) > 1) {
        "More than one race"
    } else {
        races[w]
    }
})
demo$Race <- factor(
    demo$Race,
    levels = c(
        "White",
        "Black or African American",
        "American Indian or Alaska Native",
        "Asian",
        "More than one race",
        "Native Hawaiian or Other Pacific Islander",
        "Some other race"
    ),
    labels = c(
        "White",
        "Black or African American",
        "Other race",
        "Other race",
        "More than one race",
        "Other race",
        "Other race"
    )
)
demo$Hispanicrace <- factor(
    demo$Hispanicrace,
    levels = 1:2,
    labels = c("Yes", "No")
)
# F508
demo$F508 <- factor(
    demo$F508,
    levels = 1:3,
    labels = c("F508del/F508del", "F508del/Other", "Other/Other")
)
# Age at year end
annual <- read.csv(
    "./Data_Raw/CFF19_Annualized_Del1.csv",
    na.strings = c("", "N/A")
)
annual <- annual %>% rename(reviewyear = ReviewYear)
# Merge
encounter <- left_join(
    encounter,
    demo %>% select(eDWID, Race, Hispanicrace, Gender, F508, MutClass),
    by = join_by(eDWID)
)
encounter <- left_join(
    encounter,
    annual %>% select(eDWID, reviewyear, Age_YrEnd),
    by = join_by(eDWID, reviewyear)
)
# Limit to those with two hour OGTT and HbA1c at the same encounter
# Also remove years before 2010 and those under age 10
roc_df <- encounter %>%
    arrange(eDWID, reviewyear, encounternum) %>%
    select(
        eDWID,
        reviewyear,
        Age_YrEnd,
        Race,
        Hispanicrace,
        Gender,
        F508,
        MutClass,
        twohour_bloodglucose,
        hgba1c
    ) %>%
    filter(
        reviewyear >= 2010,
        Age_YrEnd >= 10,
        hgba1c >= 3.5 & hgba1c <= 15
    ) %>%
    drop_na() %>%
    group_by(eDWID) %>%
    filter(row_number() == n()) %>%
    ungroup() %>%
    mutate(
        `OGTT 2-hour` = cut(
            twohour_bloodglucose,
            breaks = c(-Inf, 140, 200, Inf),
            labels = c("<140", "[140,200)", ">=200")
        ),
        `Age Group` = cut(Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
    )
# Metrics to report
metrics <- c(
    "thresh",
    "specificity",
    "sensitivity",
    "accuracy",
    "tn",
    "tp",
    "fn",
    "fp",
    "npv",
    "ppv"
)
```
